---
title: "Vaccine Data"
output:
  html_document:
    df_print: paged
---

It makes me happy to see the vaccine numbers tick up every day, and I also wanted to see how local jurisdictions are doing relative to the country.

'Eld_pop_share' is the percent of the adult population that is age 65+.

Dose_1_share and Dose_2_share are percentages of the adult population, and Dose_1_share_eld and Dose_2_share_eld are percentages of the 65+ population. 

You can see that areas that look to be behind the national average of adult population share look more reasonable when looking at the elderly share.This makes sense as the vaccine distributions were probably (hopefully?) more in line with elderly population than total population.

```{css, echo=FALSE}
        tr:hover {
          background-color: #ffff99;
        }
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)

# Get the latest VA file
va <- read_csv("https://data.virginia.gov/api/views/28k2-x2rj/rows.csv?accessType=DOWNLOAD")
head(va)
colnames(va)

# Get the total doses by county
va_totals <- 
  va %>% group_by(FIPS, Locality, `Dose Number`) %>%  
      summarise(total = sum(`Vaccine Doses Administered Count`))
va_totals <- va_totals %>% 
  pivot_wider(id_cols = c(FIPS, Locality), values_from = total, names_from = `Dose Number`, 
              names_prefix = 'Dose_')

# Get 2019 ACS 5-year population by age
# https://data.census.gov/cedsci/table?q=population&g=0100000US.050000&tid=ACSDP5Y2019.DP05&hidePreview=true
pop <- read_csv(file = "ACSDP5Y2019.DP05_data_with_overlays_2021-03-04T082446.csv")
pop2 <- pop %>% 
        mutate(FIPS = substr(GEO_ID,10,14), 
               DP05_0021E = as.numeric(DP05_0021E),   
               DP05_0024E = as.numeric(DP05_0024E),
               DP05_0001E = as.numeric(DP05_0001E)) %>% 
        select(FIPS, DP05_0021E, DP05_0024E, DP05_0001E)

# Drop the secondary column names
pop2 <- pop2[-1,]

# Add the pop to vaccines
va_with_pop <- inner_join(va_totals, pop2, by = "FIPS")

# Calculate vaccines as share of pop
va_with_pop2 <- mutate(va_with_pop,
                       Eld_pop_share = DP05_0024E / DP05_0021E,
                       Eld_pop_share = scales::percent(Eld_pop_share, accuracy = 0.01),
                       Dose_1_share = Dose_1 / DP05_0021E, 
                       Dose_2_share = Dose_2 / DP05_0021E,
                       Dose_1_share_eld = Dose_1 / DP05_0024E, 
                       Dose_2_share_eld = Dose_2 / DP05_0024E,
                       Dose_1_share = scales::percent(Dose_1_share, accuracy = 0.01),
                       Dose_2_share = scales::percent(Dose_2_share, accuracy = 0.01),
                       Dose_1_share_eld = scales::percent(Dose_1_share_eld, accuracy = 0.01),
                       Dose_2_share_eld = scales::percent(Dose_2_share_eld, accuracy = 0.01))
                       

# List of NOVA counties per wikipedia
nova <- c("51510","51013","51043","51047","51059","51061","51069","51107",
          "51113","51153","51157","51177","51179","51187","51600","51610",
          "51630","51683","51685","51840")

nova_vaccines <- va_with_pop2 %>% 
  filter(FIPS %in% nova)

nova_total <- nova_vaccines %>% 
  ungroup() %>% 
  summarise(Total_Dose_1 = sum(Dose_1), Total_Dose_2 = sum(Dose_2), pop = sum(DP05_0021E),
            eld_pop = sum(DP05_0024E)) %>% 
  mutate(Dose_1_share = Total_Dose_1 / pop, 
         Dose_2_share = Total_Dose_2 / pop,
         Dose_1_share_eld = Total_Dose_1 / eld_pop, 
         Dose_2_share_eld = Total_Dose_2 / eld_pop,
         Dose_1_share = scales::percent(Dose_1_share, accuracy = 0.01),
         Dose_2_share = scales::percent(Dose_2_share, accuracy = 0.01),
         Dose_1_share_eld = scales::percent(Dose_1_share_eld, accuracy = 0.01),
         Dose_2_share_eld = scales::percent(Dose_2_share_eld, accuracy = 0.01),
         Eld_pop_share = eld_pop / pop,
         Eld_pop_share = scales::percent(Eld_pop_share, accuracy = 0.01))

va_total <- va_with_pop2 %>% 
  ungroup() %>% 
  summarise(Total_Dose_1 = sum(Dose_1), Total_Dose_2 = sum(Dose_2), pop = sum(DP05_0021E),
            eld_pop = sum(DP05_0024E)) %>% 
  mutate(Dose_1_share = Total_Dose_1 / pop, 
         Dose_2_share = Total_Dose_2 / pop,
         Dose_1_share_eld = Total_Dose_1 / eld_pop, 
         Dose_2_share_eld = Total_Dose_2 / eld_pop,
         Dose_1_share = scales::percent(Dose_1_share, accuracy = 0.01),
         Dose_2_share = scales::percent(Dose_2_share, accuracy = 0.01),
         Dose_1_share_eld = scales::percent(Dose_1_share_eld, accuracy = 0.01),
         Dose_2_share_eld = scales::percent(Dose_2_share_eld, accuracy = 0.01),
         Eld_pop_share = eld_pop / pop,
         Eld_pop_share = scales::percent(Eld_pop_share, accuracy = 0.01)
         )

va_total2 <- va_total %>% 
  mutate(Locality = "VA Total",
         Dose_1 = Total_Dose_1,
         Dose_2 = Total_Dose_2) %>% 
  select(-pop, -eld_pop, -Total_Dose_1, -Total_Dose_2)

# Get US Vaccine totals
us_total <- read_csv("https://raw.githubusercontent.com/youyanggu/covid19-cdc-vaccination-data/main/cdc_vaccination_trends_data.csv")
us_total  

us_raw <- jsonlite::fromJSON("https://covid.cdc.gov/covid-data-tracker/COVIDData/getAjaxData?id=vaccination_trends_data")
us_total <- as_tibble(us_raw[["vaccination_trends_data"]])
current_total <- us_total[1,]
current_total
data_date <- as.character(us_total[1,1])
current_total <- select(current_total, Date, Administered_Cumulative, Admin_Dose_1_Cumulative,
                        Admin_Dose_2_Cumulative)
total_pop <- pop2 %>% 
  summarise(total_adult = sum(DP05_0021E), total_eld = sum(DP05_0024E))

current_total <- cbind(current_total,total_pop)
current_total2 <- current_total %>% 
  transmute(
         Locality = "US",
         Eld_pop_share = total_eld / total_adult,
         Eld_pop_share = scales::percent(Eld_pop_share, accuracy = 0.01),
         Dose_1 = Admin_Dose_1_Cumulative,
         Dose_2 = Admin_Dose_2_Cumulative,
         Dose_1_share = Admin_Dose_1_Cumulative/total_adult, 
         Dose_1_share = scales::percent(Dose_1_share, accuracy = 0.01),
         Dose_1_share_eld = Admin_Dose_1_Cumulative/total_eld, 
         Dose_1_share_eld = scales::percent(Dose_1_share_eld, accuracy = 0.01),
         Dose_2_share = Admin_Dose_2_Cumulative/total_adult, 
         Dose_2_share = scales::percent(Dose_2_share, accuracy = 0.01),
         Dose_2_share_eld = Admin_Dose_2_Cumulative/total_eld, 
         Dose_2_share_eld = scales::percent(Dose_2_share_eld, accuracy = 0.01)
         )
current_total2
nova_vaccines2 <- nova_vaccines %>% 
  ungroup() %>% 
  select(Locality, Eld_pop_share, Dose_1, Dose_2, Dose_1_share, Dose_2_share, Dose_1_share_eld, Dose_2_share_eld)

nova_total2 <- nova_total %>% 
  ungroup() %>% 
  mutate(Locality = "NoVA Total", Dose_1 = Total_Dose_1, Dose_2 = Total_Dose_2) %>% 
  select(Locality, Eld_pop_share, Dose_1, Dose_2, Dose_1_share, Dose_2_share, Dose_1_share_eld, Dose_2_share_eld) 
  
out1 <- rbind(current_total2, va_total2, nova_total2, nova_vaccines2)
out1 <- mutate(out1, Dose_1 = scales::comma(Dose_1),
               Dose_2 = scales::comma(Dose_2))

weekly_average <- us_total %>% 
  filter(Location == "US" & date_type == "Report") %>% 
  arrange(desc(Date))
weekly_average <- weekly_average[1,]
weekly_average <- as.numeric(weekly_average$Administered_daily_change_report_7dayroll)
weekly_average
remaining_doses <- 2 * (total_pop[1,1] - current_total2$Dose_1) + (current_total2$Dose_1
                                                                - current_total2$Dose_2)
as.numeric(remaining_doses)
days <- as.numeric(remaining_doses/weekly_average)
end_date <- Sys.Date() + days

```
The data are as of: **`r data_date`**.

The current 7-day average doses per day is: **`r scales::comma(weekly_average)`**.

At this rate, all adults could be vaccinated by: **`r end_date`**.

```{r, echo=FALSE}
knitr::kable(out1)
```

[Github](https://github.com/abibler/DMV_Covid_Vaccine_Tracking)
